### 2019-08-01 Thesis/pages/design.tex
Changes; - Added a lot of content to the design stage of the report

The attack on Trustwords involves generating "near-collision" keys. These are keys that are deemed a match by the phonetic similar metric (More on this aspect in the following sections). A similarity metric is an algorithm designed to determine if two words are phonetically a match. For example, the words "THEIR" and "THERE" will be phonetically be a match whereas the words "DARK" and "PRINCIPLE" are not phonetically matching.

@@

\subsection{Similarity metrics}

\subsubsection{Soundex}
One of the first requirements for the attack is quantifying "phonetic similarity". There are many algorithms currently available to provide this function, is known as Soundex. It is one of the most famous example of a phonetic algorithm due to its implementation into major database clients like MySQL\cite{mysql_soundex}, Oracle\cite{moved_2005} and PostgreSQL\cite{postgresql}. Originally designed for indexing names by sound alongside spelling mistakes due to small changes in letters. However, due to its design being based on the phonetic features it is highly suitable for this use-case.

Soundex produces a four digit code for each word assessed.
The first letter of the word is retained alongside the removal of all of a, e, i, o, u, y, h and w. The remaining letters are then mapped to numbers.

@@

    \caption{Soundex mappings of letters to numbers}
\end{figure}

Further steps are performed on the code, as the technical details do not add anything to the discussion, the steps of the algorithm can be found in Appendix TODO.

\subsubsection{Soundex Issues}

@@

    \item \textbf{Silent consonants.} Soundex does not have logic embedded to deal with silent consonants within words.

    \item \textbf{Poor precision.} Due to the previously discussed point of a small code resulting in lack of precision. \cite{patman2001soundex} re-iterates this point but in the context of name matching where Soundex's poor performance was demonstrated. Soundex only gained and overall accuracy 36.37\% when matching names within a database.
\end{enumerate}

@@

\subsubsection{NYSIIS}
The New York State Identification and Intelligence System (NYSIIS) phonetic code was created for use matching the phonetics of American names. It was created due to the presence of hispanic names in the American based databases (this was an aspect Soundex was known to have low accuracy with). However, due to it having embedded rules to handle word phonetics it would, again be applicable in this application.

@@

\subsubsection{Metaphone}
Metaphone was invented by Lawrence Philips in 1990\cite{philips1990hanging} in response to the deficienies in Soundex. It improves on Soundex by including information around inconsistency and variation in English spelling in an attempt to create a more accurate phonetic representation. Metaphone is arguably on the same level of ubiquity as that of Soundex with it finding itself implemented in languages such as big examples like PHP\cite{php}. Further work would involve the implementation of newer versions of Metaphone. Double Metaphone (2000) and Metaphone 3 (2009) claim to improve over the original version due to further research performed by Philips. The original version was chosen due to its historical wide spread usage alongside the case of Metaphone 3 being proprietary code and, therefore, under a paid licence.

\subsubsection{Levenshtien Distance}
Levenshtein distance is a string metric designed to measure the 'distance' between two strings. It is simply the number of single-character edits (insertions, deletions or substitutions) required to reach the other string. Edit distance is not technically designed as a phonetic algorithm, but due to similar-sounding words often being spelt in similar ways\cite{hettiarachchi2012sparcl} Levenshtien distance was deemed another suitable metric.

@@

\subsubsection{Phonetic Vectors}
Phonetic Vectors is the unique addition to the chosen set. Created by Allison Parrish in 2017\cite{parrish2017poetic}, Phonetic vectors is as the name suggests the vectorization of a words phonetics. This allows a words phonetics to be represented using a continuous number in vector space.

Phonetic features are used in this work as a way to compare the similarity of the phonemes. Extensive prior work has gone into producing models of features that map to phonemes. \cite{chomsky1968sound}\cite{ladefoged1969measurement}\cite{bradlow2010perceptual}. Features, therefore, a newer attempt at mapping the varying and inconsistent rules around pronunciation of the English language. The vectors are created using lists phonemes from the CMU Pronouncing Dictionary. Phonemes are the phonetic elements that construct a word. For example the word "RING" translated into the phonemes \verb|/R IH NG/|. Using all words available and mapping all phonemes to phonetic features a list of unique features is created per-word.

@@

\textbf{TODO: } Explain why being vectors is useful and its appliactions

\textbf{TODO: } Do a sample conversions table

\subsection{Alternative Similarity Metrics}

\subsubsection{Match Rating Approach}
Matching Rating Approach (MRA) is another algorithm designed to match names within a database, therefore, its operation can be grouped with that of NYSIIS and Soundex. MRA was discarded as an option. This is because of the lack of known utilization in any substantial real world use-cases alongside its similarity to more well established algorithms of Soundex and NYSIIS.

@@

\subsubsection{Caverphone}
Another notable alternative solution is that of Caverphone that was designed in New Zealand. Caverphone as is the case with the vast majority of phonetic algorithms was designed for use with name matching. Caverphone was not chosen to similar reasons to that of MRA alongside its optimization for accents in location of New Zealand it was conceived in. Therefore, the low level of utilization alongside the unique design features that suggest it may not be suited for this application. This has, however, not be assessed empirically and thus would be a candidate for further work.

% Other alternatives:
% - Match Rating Approach
%   - Couldn't find an algorithm??
%
% - Caverphone

\textbf{TODO: } Need to justify why

\section{Design of GreenOnion}
And the way it works. Need to explain it is based on Scallion but improved with the use of C++ alongside a bloom filter.

>>>>

The attack on Trustwords involves generating "near-collision" keys. 
This will attempt to provide the base to answer Research Questions \ref{goal:numberOfTrustwords}, \ref{goal:complexity} and \ref{goal:hardwareRequired}.

Near-collision keys are keys that are deemed a match by the phonetic similar metric (More on this aspect in the following sections). A similarity metric is an algorithm designed to determine if two words are phonetically a match. For example, the words "THEIR" and "THERE" will be phonetically be a match whereas the words "DARK" and "PRINCIPLE" are not phonetically matching.

@@

\section{Similarity metrics}
\label{sec:metrics}
One of the first requirements for the attack is quantifying "phonetic similarity". This section is looking to provide an answer for the Research Question \ref{goal:phoneticSimilarity}. There are many algorithms currently available to provide this functionality. This section will describe algorithms alongside the reasons they were selected for assessment later in the project.

\subsection{Soundex}
One of the earliest example of a phonetic algorithm is known as Soundex. It is one of the most famous algorithm due in part to its implementation into major database clients like MySQL\cite{mysql_soundex}, Oracle\cite{moved_2005} and PostgreSQL\cite{postgresql}. Originally designed for indexing names by phonetics alongside spelling mistakes due to transpositions in letters. Therefore, due to it being based on the phonetic features, it is highly suitable for this project's use-case.

Soundex produces a four digit code for each word assessed.
The first letter of the word is retained alongside the removal of all of a, e, i, o, u, y, h and w. The remaining letters are then mapped to numbers\footnote{Further steps are performed on the code, as the technical details do not add anything to the discussion, the steps of the algorithm can be found in Appendix TODO.}. These mappings are displayed in Figure \ref{fig:soundexMap}.

@@

    \caption{Soundex mappings of letters to numbers}
    \label{fig:soundexMap}
\end{figure}

\subsection{Soundex Issues}
\label{sec:soundexIssues}

@@

    \item \textbf{Silent consonants.} Soundex does not have logic embedded to deal with silent consonants.

    \item \textbf{Poor precision.} Due to the previously discussed point of a small code space. \cite{patman2001soundex} re-iterates this point but in the context of name matching where Soundex's poor performance was demonstrated. Soundex only gained and overall accuracy 36.37\% when matching names within a provided database.
\end{enumerate}

@@

\subsection{NYSIIS}
The New York State Identification and Intelligence System (NYSIIS) phonetic code was created for use matching the phonetics of American names. It was created due to the presence of hispanic names in the American based databases (this was an aspect Soundex was known to have low accuracy with). However, due to it having embedded rules to handle word phonetics it would, again be applicable in this application.

@@

\subsection{Metaphone}
Metaphone was invented by Lawrence Philips in 1990\cite{philips1990hanging} in response to the deficiencies in Soundex. It improves on Soundex by including information around inconsistency and variation in English spelling in an attempt to create a more accurate phonetic representation. Metaphone is arguably on the same level of ubiquity as that of Soundex with it finding itself implemented in languages such as PHP\cite{php}.

Further work would involve the implementation of newer versions of Metaphone. Double Metaphone (2000) and Metaphone 3 (2009) claim to improve over the original version due to further research performed by Philips. The original version was chosen due to its historical widespread usage.

\subsection{Levenshtien Distance}
Levenshtein distance is a string metric designed to measure the 'distance' between two strings. It is simply the number of single-character edits (insertions, deletions or substitutions) required to reach the other string. Edit distance is not technically designed as a phonetic algorithm, but due to similar-sounding words often being spelt in similar ways\cite{hettiarachchi2012sparcl} Levenshtien distance was deemed another suitable metric.

@@

\subsection{Phonetic Vectors}
Phonetic Vectors is the unique addition to the chosen set. Created by Allison Parrish in 2017\cite{parrish2017poetic}, Phonetic vectors is as the name suggests the vectorization of a words phonetics. This allows a words phonetics to be represented in vector space.

Phonetic features are used in this work as a way to compare the similarity of a words phonemes. Phonemes are the phonetic elements that construct a word. For example the word "RING" translated into the phonemes \verb|/R IH NG/|. Extensive prior work has gone into producing models of features that map to phonemes. \cite{chomsky1968sound}\cite{ladefoged1969measurement}\cite{bradlow2010perceptual}. Features, therefore, are an attempt at mapping the varying and inconsistent rules around pronunciation of the English language. The vectors are created using lists phonemes from the CMU Pronouncing Dictionary. Using all words available in the CMU Dictionary and mapping all phonemes to phonetic features produces a list of unique features.

@@

The ability to perform operations and measure dissimilarity allows for a plethora of applications. One possible application of note would be the creation of a wordlist where the phonetic difference (vector distance) is maximized. This, if the vector mappings are an accurate representation could allow a creation of a phonetically distinctive wordlist. This will be discussed further at a later stage of the report.

\section{Alternative Similarity Metrics}

\subsection{Match Rating Approach}
Matching Rating Approach (MRA) is another algorithm designed to match names within a database, therefore, its operation can be grouped with that of NYSIIS and Soundex. MRA was discarded as an option. This is because of the lack of known utilization in any substantial real world use-cases alongside its similarity to more well established algorithms of Soundex and NYSIIS.

@@

\subsection{Caverphone}
Another notable alternative solution is that of Caverphone that was designed in New Zealand. Caverphone as is the case with the vast majority of phonetic algorithms was designed for use with name matching. Caverphone was not chosen to similar reasons to that of MRA alongside its optimization for accents in location of New Zealand it was conceived in. Therefore, the low level of utilization alongside the unique design features that suggest it may not be suited for this application. This has, however, not be assessed empirically and thus would be a candidate for further work.

\begin{table}[h!]
    \centering
    \begin{tabular}{ll}
        Metric & Output \\
        \hline    
        Soundex & T614 \\
        NYSIIS & TRAVAL\\
        Metaphone & TRFL\\
        Caverphone & TRF111\\
        MRA & TRVL
    \end{tabular}
    \caption{The various phonetic encodings of the word "Travel"}
\end{table}

\section{Design of GreenOnion}
To answer Research Questions \ref{goal:complexity} and \ref{goal:hardwareRequired} fully, it is necessary to implement an actual code base that will be used to generate actual keys.

Inspiration of the design of this tool was taken from a tool called Scallion\footnote{\url{https://github.com/lachesis/scallion}}. Scallion was designed by Richard Klafter and Eric Swanson and used to demonstrate that 32-bit PGP key ids were insufficient. Keeping with the onion-based theme the proposed tool is known as GreenOnion and is a re-write of the tools structure in C++. This language was chosen due to the well understood efficiency benefits. The proposed tool differs from Scallion most notably in its ability to concurrently search for a large number of keys, GreenOnion improves on this substantially. More implementation and experimental details will be discussed in later chapters.

The tool should take two keys as parameters (Uncontrolled/Controlled) and a chosen similarity metric and produce a list of target keys fingerprints. This list is then used as a search criteria when hashing a checking a large number of keys. In order to utilise the high parallel nature of the GPU to compute the hash of a large number of keys the tool will utilise a GPGPU (General-purpose computing on graphics processing units) framework. The chosen framework was OpenCL due to its support for the chosen language (C++) and platform (Linux). OpenCL allows the creation of code chunks refereed to as "kernels" to be executed concurrently, this will provide a massive speed increase compare to the sequential nature of the CPU. Technical details will be explained further in Chapter \ref{cha:Implementation}.

\section{Experiment Design}
In order to answer some of the research questions empirical evidence is required and, thus, experiments are required. In this section the designs and considerations of the chosen experiments will be discussed.

\subsection{Metric performance}
In order to answer Research Question \ref{goal:bestMetrics} and reduce the number of metrics to asses in the later rounds, the performance of similarity metrics required comparison. The thinning out of metrics is required due to limited resources, further work could involve repeating this work with a much more varied selection of metrics. As explained in a previous section (See Section \ref{sec:metrics}) the chosen metrics up for assessment are Soundex, Metaphone, NYSIIS, Levenshtein and Phonetic vectors.

The design for the experiment involves assessing the quality of the metrics matches by matching them to human perception of similarity.

\subsubsection{Matches}
A match for each code based metric (Soundex, Metaphone and NYSIIS) is when the codes are identical. However, for the other cases where the difference is variable other ways of defining a match are required.

In the case of Levenshtein values of similarity are discrete due to it being the number of single character edits. Therefore, this requires less deliberation. Match size was used as a way to decide on a cut-off point.

\begin{table}[h!]
    \centering
    \begin{tabular}{ll}
        Metric & Matches \\
        \hline
        Soundex     & 1,527,554 \\
        Metaphone   & 412,916 \\
        NYSIIS      & 188,474 \\
        \hline
        Levenshtein (Distance 1) [L1] & \textbf{97,730} \\
        Levenshtein (Distance 2) [L2] & 1,070,656
    \end{tabular}
    \caption{Levenshtein number of matches comparison}
    \label{tab:matchSize}
\end{table}

As it can be seen from Table \ref{tab:matchSize} the number of matches for L2 is massively larger than all but Soundex. As the issues with Soundex have been discussed in previous sections (See Section \ref{sec:soundexIssues}) it can be excluded as an abnormally in this context as it is only included due to its historical significance. Therefore, due to the excessively high value for a L2, L1 was chosen as the Levenshtein cap for defining a match.

Match size was also used to define the cap for the phonetic vectors. This is less distinct that that of Levenshtein due to the continuous nature of the distance between two vectors, however the complexity was reduced by limiting it to increments of 0.5.

\begin{table}[h!]
    \centering
    \begin{tabular}{ll}
        Metric & Matches \\
        \hline
        Soundex     & 1,527,554 \\
        Metaphone   & 412,916 \\
        NYSIIS      & 188,474 \\
        Levenshtein & 97,730 \\
        \hline
        Phonetic Vector (Distance 3.0)  & 14,550 \\
        Phonetic Vector (Distance 4.0)  & 73,962 \\
        Phonetic Vector (Distance 4.5)  & \textbf{216,156} \\
        Phonetic Vector (Distance 5.0)  & 685,516 \\
    \end{tabular}
    \caption{Phonetic vector number of matches comparison}
    \label{tab:vectorMatchSize}
\end{table}

Table \ref{tab:vectorMatchSize} contains the match size comparison for Phonetic vectors distance. Distance 4.5 was chosen as the suitable size due to the total matches fitting well within the other metrics. Distance 4.0 was the other alternative, this was discarded due to its small size. Any metric with a number of matches below 100,000 results in an inadequate number of possible near-collision keys later in the process, a lower number of matches may imply better quality but a balance is required between computational cost and attack quality. This will be discussed in more detail in later chapters.
